---
title: "9.26.17 HW"
author: "Christopher Law"
date: "September 23, 2017"
output:
  pdf_document: default
  html_document: default
---

# Reflection on readings

  On a practice GRE recently, I got a question that required me to interpret a chart with two vertical axis. I starred at it for a minute or so and couldn't figure out which axis corresponded to one line. It was maddening! The points on the two lines used two different symbols but I couldn't interpret which line mapped to which axis. I totally agree with the authors of this week's reading that these types of charts are confusing and a sign of naïve data analysis.  
  The blog post from junk charts that poked holes it the NYT chart was interesting in that a lay person could not have noticed the poor representation of data; the author could read between the lines and see problems only because he/she had a strong statistical background. This makes me think that the creator of the chart may have been using deception intentionally because he/she knew they could get away with it. Pretty dangerous stuff!  
  I loved the awesome YouTube video. What a cool way of representing a large set of complicated data. Not only was the presentation impressive but the story was powerful and insightful. I think it's a great example - maybe the best we've analyzed yet - of effective data visualization.   
  I thought the piece on perceptualedge was well-written. I enjoyed the use of storytelling employed by the author so I could understand how conclusions were reached. This is powerful because I can undergo the same thought exercise and make similar conclusions - thus strengthening my agreeance with the points. The logic for bar chars seems especially sound. They are meant to compare quantities and presenting data on different scales manipulates the truth behind the data. I also thought it was interesting how people assume that there's meaning behind the intersection of lines but this is fallacious reasoning if the scales are different (and even sometimes if they are the same).  
  Overall, I like the idea of using two different charts. They are much easier to read than the complicated alternative presented by Stephanie Evergreen. I think her chart is difficult because the optimal direction for the lines is to the right and down. However, this is different from most other charts that we see where the top-right is the ideal location. I think this is manipulative to play with people's default interpretations. It reminds me of the much more egregious data visualization presented in class about the murder rate (or something similar) in Florida. One of the main purposes of a chart is to get people to quickly interpret information. This is precisely why I think it's untruthful to mess with people's default interpretations.  
  The Cairo chapter was a good refresher of what I'm learning in my Stats class. I can see why some universities require all their students to take a statistics class. I guess my bigger fear is if you're trying to communicate with a group uneducated in stats.  



# Hans Rosling redux

```{r load-packages-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(scales)
library(gapminder)
library(ggthemes)

# Load the gapminder data from the gapminder package with data()
data(gapminder)

# Look at the variables in the data frame
glimpse(gapminder)

# Create a data frame with only rows from 1997
gapminder_1997 <- gapminder %>%
  filter(year == 1997)
```


## Univariate analysis

Plot a histogram of life expectancy in 1997 (`lifeExp`). Choose an appropriate number of bins. (You get some code to help start you off):

```{r lifeexp-hist}
ggplot(gapminder_1997, aes(x = lifeExp)) +
  geom_histogram(bins = 10) +
  ggthemes::theme_economist()
```

Plot a density plot of life expectancy in 1997. Fill it with some color so it doesn't look sad and empty.

```{r lifeexp-density}
ggplot(gapminder_1997, aes(x = lifeExp)) +
  geom_density(fill = "orange3")
```

Plot a histogram of GDP per capita in 1997 (`gdpPercap`).

```{r gdp-hist}
ggplot(gapminder_1997, aes(x = gdpPercap)) +
  geom_histogram(bins = 10) 
```

Plot a density plot of GDP per capita in 1997.

```{r gdp-density}
ggplot(gapminder_1997, aes(x = gdpPercap)) +
  geom_density(fill = "yellow")


```

## Univariate analysis by groups

Plot multiple violin plots of GDP per capita in 1997 by continent. Fill the continents with different colors. Add points at 50% transparency (hint: `alpha = 0.5` in a `geom_point()` layer would help). (You get some code to help start you off):

```{r gdp-continent}
ggplot(gapminder_1997, aes(x = continent, y = gdpPercap, fill = continent)) +
  geom_violin() +
  geom_point(alpha = .5) 

```

Plot multiple boxplots (hint: `geom_boxplot()`) of GDP per capita in 1997 by continent. Fill the continents with different colors.

```{r gdp-continent-box}
ggplot(gapminder_1997, aes(x = continent, y = gdpPercap, fill = continent)) +
  geom_boxplot()

```

Plot multiple violin plots of life expectancy in 1997 by continent, also with filled continents and semi-transparent points (hint: do basically what you did above, but with `lifeExp` instead of `gdpPercap`)

```{r lifeexp-continent}
ggplot(gapminder_1997, aes(x = continent, y = lifeExp, fill = continent)) +
  geom_violin() + 
  geom_point(alpha = .5)
```

Plot overlapping density plots of life expectancy in 1997 across continents. Oceania has very few observations, so omit it from the data (I create a filtered data frame for you below). Fill each continent with a color and make each density plot 50% transparent

```{r continent-densities}
gapminder_1997_sans_oceania <- gapminder_1997 %>%
  filter(continent != "Oceania")

ggplot(gapminder_1997_sans_oceania, aes(x = lifeExp, fill = continent)) + 
  geom_density(alpha = 0.5)
```

## Bivariate analysis

Plot health (`lifeExp`) vs. wealth (`gdpPercap`) in 1997. Color each point by continent. (You get some code to help start you off):

```{r health-wealth-basic}
ggplot(gapminder_1997, aes(x = gdpPercap, y = lifeExp, color = continent)) +
  geom_point()
```

Make that same plot, but add `coord_trans(x = "log10")` as a layer.

```{r health-wealth-transformed}
ggplot(gapminder_1997, aes(x = gdpPercap, y = lifeExp, color = continent)) +
  geom_point() +
  coord_trans(x = "log10")
```

What's different?  
_The axis is scaled via a logarithm_

Plot health vs. wealth again (without a logged x-axis), and add a `geom_smooth()` layer.

```{r health-wealth-smooth, message=FALSE, warning=FALSE}
ggplot(gapminder_1997, aes(x = gdpPercap, y = lifeExp, color = continent)) +
  geom_point() +
  geom_smooth()
```

By default, R will choose `method = "loess"` to plot the line. What is "loess"? (hint: see pages 240-41 in Cairo). Change the smoothing method to `method = "lm"` (`lm` here stands for "linear model")

```{r health-wealth-lm}
ggplot(gapminder_1997, aes(x = gdpPercap, y = lifeExp, color = continent)) +
  geom_point() +
  geom_smooth(method = "lm")
```

What's different?

Plot health vs. wealth *with* a logged x-axis *and* with a loess smooth.

```{r health-wealth-log-loess, message=FALSE, warning=FALSE}
ggplot(gapminder_1997, aes(x = gdpPercap, y = lifeExp, color = continent)) +
  geom_point() +
  coord_trans(x = "log10") +
  geom_smooth()
```

Plot health vs. wealth *with* a logged x-axis *and* a linear smooth (`lm`). 

```{r health-wealth-log-lm}
ggplot(gapminder_1997, aes(x = gdpPercap, y = lifeExp, color = continent)) +
  geom_point() +
  coord_trans(x = "log10") +
  geom_smooth(method = "lm")
  
```


## Fancy stuff

Here's a fancy, production-quality version of the health-wealth plot. Explain what each of these layers are doing:

- `ggplot(gapminder_1997, aes(...))`: Defines the data set and the main aesthetics (including defining variables, size of the points, and color of the points)
- `geom_point()`: specifying we want a scatter plot
- `guides()`: adjusts the legend
- `labs()`: specifying the format of the labels
- `scale_x_continuous()`: providing the scale for the x-axis
- `coord_trans()`: scales the x-axis via logarithm
- `theme_light()`: changes the visual theme of the graph
- `theme()`: specifies the position of the legend and the grid lines

```{r health-wealth-fancy}
ggplot(gapminder_1997, aes(x = gdpPercap, y = lifeExp, 
                           size = pop, color = continent)) +
  geom_point() + 
  guides(size = FALSE, color = guide_legend(title = NULL)) +
  labs(x = "GDP per capita", y = "Life expectancy") +
  scale_x_continuous(labels = dollar) +
  coord_trans(x = "log10") + 
  theme_light() + 
  theme(legend.position = "bottom",
        panel.grid.minor.x = element_blank())
```


## Multiple years

Look at the relationship between health and wealth in 1992, 1997, 2002, and 2007 all in one plot. To do this, you can't use the `gapminder_1997` data frame anymore, since that's just 1997â???"you'll need to create a new data frame. 

Color each point by continent and resize each point by population. Place each of the four years in a separate facet (hint: look at the documentation for `facet_wrap()`). (You get some code to help start you off):

```{r health-wealth-post-1992}
gapminder_after_1992 <- gapminder %>%
  filter(year >= 1992)

ggplot(gapminder_after_1992, aes(x = gdpPercap, y = lifeExp, color = continent, size = pop)) +
  geom_point() + 
  coord_trans(x = "log10") +
  facet_wrap(~year)
```

Create a similar plot to show the relationship between health and wealth in 1952, 1957, 1962, and 1967. Again, you won't be able to use either the `gapminder_1997` or the `gapminder_after_1992` data framesâ???"you'll have to create a new data frame. And I won't give you code for that.

```{r health-wealth-early-cold-war}
gapminder_52thru67 <- gapminder %>% 
  filter(year == 1952 |year == 1957 |year == 1962 |year == 1967)

ggplot(gapminder_52thru67, aes(x = gdpPercap, y = lifeExp, color = continent, size = pop))+
  geom_point() +
  coord_trans(x = "log10") +
  facet_wrap(~year)

```